Face Recognition: Too Bias, or Not Too Bias?

### 2020-04-13 cvprworkshop.tex
moved back into section (ie organized by file)

% \usepackage{FG2019}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage[acronym]{glossaries}
\usepackage{acronym}

\newcommand{\minitab}[2][c]{\begin{tabular}{#1}#2\end{tabular}}

@@

% \let\vec\mathbf

% \cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{0002} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

@@

@@

%%%%%%%%% ABSTRACT
\begin{abstract}
We reveal critical insights into problems of bias in state-of-the-art \gls{fr} systems using a novel \gls{bfw} dataset: data balanced for gender and ethnic groups. We show variations in the optimal scoring threshold for face-pairs across different subgroups. Thus, the conventional approach of learning a  global threshold for all pairs resulting in performance gaps among subgroups. By learning subgroup-specific thresholds, we not only mitigate problems in performance gaps but also show a notable boost in the overall performance. Furthermore, we do a human evaluation to measure the bias in humans, which supports the hypothesis that such a bias exists in human perception. For the \gls{bfw} database, source code, and more, visit \href{https://github.com/visionjo/facerec-bias-bfw}{github.com/visionjo/facerec-bias-bfw}
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \begin{bibunit}

\glsresetall
\section{Introduction}
    As more of society becomes integrated with \gls{ml}, bias, fairness, and the formalization of \gls{ml} standards attract more attention~\cite{10.1007/978-3-030-13469-3_68, anne2018women, wang2018racial}. Thus, an effect of the growing dependency on \gls{ml} are growing concerns for biased and unfair algorithms, \eg untrustworthy and racist \gls{fr} systems~\cite{england2019,snow2018}.

    Typically, \glspl{cnn} are trained on faces identified by a detection system. Specifically, for \gls{fr}, the goal is to encode faces in an N-dimensional space such that samples of the same identity are neighbors and those of different identity of further apart. Thus, the goal is to identify subjects with face encodings. We can deploy a \gls{cnn} to encode faces, and then compare via a similarity score: assume pairs with a high enough score are \emph{genuine}; else, an \emph{imposter}.

% \begin{figure}
%     \centering
%     \includegraphics[width=\linewidth]{figures/fig1-crop.pdf}
%     \caption{Caption}
%     \label{fig:my_label}
% \end{figure}
    \begin{figure}[t!]
        \centering
        \includegraphics[width=.9\linewidth]{figures/montage.pdf}
        \caption{\textbf{\gls{bfw}.} The average face of its different subsets: \emph{top-left}: the entire \gls{bfw}; \emph{top-row} per race;  \emph{left-column}: per gender. The others represent the ethnicity and gender of the race and gender, respectively. Table~\ref{tab:ethnic-splits} defines the acronyms of subgroups.}
        \label{fig:avg-faces}
    \end{figure}

    When comparing scores, typically, a fixed threshold sets the decision boundary. Thus, features of the same identity must satisfy a criterion via a single value~\cite{deng2019arcface, liu2017sphereface, wang2018additive, wang2018cosface}. However, we found that a single (\ie global) threshold is a crude measure that leads to skewed errors - the held-out set used to determine the threshold tends to share the same distribution with the test data, which favors specific demographics that are a majority. That skew, the difference in the performance of an algorithm of certain demographics, is our definition of bias. A key question is: \emph{is \gls{fr} too biased, or not?}

\begin{table*}[!t]
    \centering
    \caption{\small{\textbf{Database stats and nomenclature, optimal thresholds ($t_o$), and accuracy scores.} \textit{Header:} Subgroup definitions. \textit{Top-row:} Statistics of \gls{bfw}. \textit{Middle:} Number of pairs for each partition. \textit{Bottom:} Accuracy with a global threshold $t_g$, the optimal threshold $t_o$, and accuracy with $t_o$ per subgroup. Specifically, $t_g$=0.259$\pm$0.002 when averaged across the five folds. Columns grouped by race and then further split by gender. Out of millions of pairs, accuracy is inconsistent across subgroups. Furthermore, $F$ tends to perform inferior to that of $M$ for all but B.}}\label{tab:ethnic-splits}
    \scalebox{.75}{
     \resizebox{\textwidth}{!}{%
    \begin{tabular}{r c c c c c c c c l}
        \toprule
        & \multicolumn{2}{c}{Asian (A)} & \multicolumn{2}{c}{Black (B)}  & \multicolumn{2}{c}{Indian (I)}& \multicolumn{2}{c}{White (W)}\\
        \cmidrule(l){2-3} \cmidrule(l){4-5} \cmidrule(l){6-7}\cmidrule(l){8-9} % spanning less than the full width of the table - you can add (r) or (l) just before the opening curly bracket to shorten the rule on the left or right side
         & Female (AF) & Male (AM) & BF & BM& IF & IM & WF & WM&Aggregated\\ % Column names row
        \midrule

       \# Faces  &  2,500&  2,500& 2,500 & 2,500& 2,500 & 2,500 & 2,500 & 2,500 &20,000 \\ % row 1
        \# Subjects & 100& 100& 100  & 100  & 100  & 100& 100 &100&800  \\ % row 2
        \# Faces / Subject  & 25 & 25    & 25 & 25 & 25  & 25  &  25 & 25 & 25\\ % row 3
        % \midrule
\specialrule{.01em}{.05em}{.05em}
            \# Positive Pairs &  30,000&  30,000& 30,000 &30,000 & 30,000 &30,000&30,000 & 30,000 &240,000 \\ % row 1
        \# Negative Pairs & 85,135&  85,232& 85,016  & 85,141  & 85,287  & 85,152& 85,223 &85,193&681,379  \\ % row 2

        \# Pairs (Total) & 115,135 & 115,232    &115016 &115,141 & 115287  & 115,152  &  115,223& 115193 & 921,379\\ % row 3
        % \midrule
        % \specialrule{.01em}{.05em}{.05em}
        % Acc$@t_g$ & 0.941 & 0.952  & 0.967 & 0.960 & 0.958 &0.962  & 0.973 & 0.981 &0.962 $\pm$0.012 \\
        % $t_o$ & 0.255 &  0.246 & 0.277 &0.242  &0.297 & 0.264& 0.233 &0.239 &0.257$\pm$0.006\\ % row 4

        % Acc$@t_o$ & 0.942 &0.952 &0.968 &0.961 &0.960 &0.962 & 0.974 &0.982 & 0.963 $\pm$ 0.012\\
        \bottomrule
    \end{tabular}}
    % \glsunset{af}
    }
    % % \glsunset{af}
    % \glsunset{am}
    % \glsunset{bf}
    % \glsunset{bm}
    % \glsunset{if}
    % \glsunset{im}
    % \glsunset{wf}
    % \glsunset{wm}
    % \vspace{-12pt}
\end{table*}

    Making matters more challenging is that race and ethnicity are loosely defined. For example, the US Census Bureau allows an individual to self-identify race.\footnote{\scriptsize\href{https://www.census.gov/mso/www/training/pdf/race-ethnicity-onepager.pdf}{www.census.gov/mso/www/training/pdf/race-ethnicity-onepager}} For this work, we define subgroups as specific sub-populations with face characteristics similar to those found in a region.

    The adverse effects of a global threshold are two-fold: \textbf{(1)} the mapping produced by a \gls{cnn} is nonuniform. Therefore, distances between pairs of faces in different demographics vary in distribution of similarity scores (Fig~\ref{fig:detection-model}); \textbf{(2)} the evaluation set is imbalanced as well. Particular demographics making up a majority of the population will carry most weight on the reported performance ratings. Reported results favor the common traits over the underrepresented. Alas, demographics like gender, ethnicity, race, and age are underrepresented in most public datasets~\cite{merler2019diversity, wang2018racial}.

    % The result is various types of biases in FR systems in favor of or against particular demographics remain a question.

    For \textbf{(1)} we propose subgroup-specific (\ie optimal) thresholds; to address \textbf{(2)}, we introduce a new benchmark for measuring bias in \gls{fr}, \gls{bfw} (Table~\ref{tab:compared} \&~\ref{tab:ethnic-splits}). \gls{bfw} allows for a fair evaluation of \gls{fr} systems, while enabling demographic-specific ratings to be reported. We use \gls{bfw} to gain a deeper understanding of the extent of bias present in \gls{soa} \gls{cnn} used to encode faces. Then, we suggest a mechanism to mitigate problems of bias with more balanced performance ratings for different demographics. Specifically, we propose using an adaptive threshold that varies depending on the characteristics of detected facial attributes (\ie gender and ethnicity, Fig~\ref{fig:avg-faces}). We show an increase in accuracy with a balanced performance for different subgroups. Similarly, we show a positive effect of adjusting the similarity threshold based on the facial features of matched faces. Thus, selective use of similarity thresholds in current \gls{soa} \gls{fr} systems provides more intuition in \gls{fr} research with a method easy to adopt in practice.

    The contributions in this paper are 3-fold: (1) Built a balanced dataset as a proxy to measure verification performance per subgroup; (2) Analyzed an unwanted bias in scores of face pairs while showing that optimal thresholds determined per subgroup balances performance; (3) Showed bias causes skewness in ratings across demographics-- the typical use of a global threshold unfavorable. We mitigate the problem via adaptive thresholds; (4) Conducted human-evaluations to demonstrate bias in human perception.\footnote{NIH-certified, \textit{Protect Humans in Research}, \textcolor{red}{IRB 19-09-08}.}
    % \begin{enumerate}
    %     \item Built a balanced dataset as a proxy to measure verification performance per subgroup.
    %     \item Analyzed an unwanted bias in scores of face pairs while showing that optimal thresholds determined per subgroup significantly boost and balance performances.
    %     \item Showed bias causes inconsistencies in ratings across demographics-- the typical use of a global threshold unfavorable. We mitigate the problem via adaptive thresholds.
    %     \item Conducted human-evaluations to demonstrate bias in human perception.\footnote{NIH-certified, \textit{Protect Humans in Research}, \textcolor{red}{IRB 19-09-08}.}
    % \end{enumerate}

    \begin{table*}[t!]

        \centering
        \caption{\textbf{\gls{bfw} compared to related datasets.} Our \gls{bfw} is balanced across ID, gender, and ethnicity (Table~\ref{tab:ethnic-splits}). Compared with \gls{dp}, \gls{bfw} provides more samples per subject and subgroups per set, while only using a single resource, VGG2. \gls{rfw}, on the other hand, supports a different task (\ie domain adaptation), and focuses on race-distribution, but not the distribution of identities.}
        \scriptsize
        \begin{tabular}{rccccccc}%x{10mm}x{3mm}x{10mm}x{8mm}}%\toprule

            \multicolumn{2}{c}{Database} & \multicolumn{3}{c}{Number of}& \multicolumn{3}{c}{Balanced Labels}\\
            \cmidrule(lr){1-2}	\cmidrule(lr){3-5} \cmidrule(lr){6-8}
            Name & Source Data & Faces &  IDs & Subgroups & ID & Ethnicity & Gender\\\midrule
            \gls{dp}~\cite{demogPairs}     & CASIA-W~\cite{yi2014learning}, VGG~\cite{schroff2015facenet} \&VGG2~\cite{Cao18} & 10,800& 600 & 6 &\checkc& \checkc &\checkc \\
            \gls{rfw}~\cite{wang2018racial}     &  MS-Celeb-1M &$\approx$80,000&$\approx$12,000& 4 & \xmark & \checkc &\xmark \\
            \gls{bfw} (ours) & VGG2 & 20,000 & 800 &8 & \checkc & \checkc &\checkc \\\bottomrule
        \end{tabular}
        \label{tab:compared}
    \end{table*}

\section{Background Information}
    \subsection{Bias in \gls{ml}}
        The progress and the commercial value of \gls{ml} is exciting; however, the trust of society in \gls{ml} will take time to build due to inherent biases. The exact definitions and implications of bias vary between sources, as do its sources and types. A common theme is that bias hinders performance ratings in ways that skew to a particular sub-population-- the source varies, whether from humans~\cite{windmann1998subconscious}, data and label types~\cite{tommasi2017deeper}, \gls{ml} models~\cite{amini2019uncovering, kim2019learning}, or evaluation protocols~\cite{stock2018convnets}. For instance, a vehicle-detection model might miss cars if training data was mostly trucks. In practice, many \gls{ml} systems learn on biased data, which could be detrimental for society.

    \subsection{Bias in \gls{fr}}
        Problems of bias in \gls{fr} have been driven by different motivations; therefore, solutions have been posed for different matters. To name a few: data augmentation~\cite{yin2019feature}, one-shot learning~\cite{ding2018one}, demographic parity and fairness with priority on privacy~\cite{huang2018generative}, domain adaptation~\cite{wang2018racial}, differences in face-based attributes across demographics~\cite{wang2018they}, and even data exploration~\cite{muthukumar2019}. Yin~\etal proposed to augment the feature space of underrepresented classes using other classes with a diverse collection of samples to encourage distributions of underrepresented classes to more closely resemble that of the others~\cite{yin2019feature}. Similarly, others formulated the imbalanced class problem as one-shot learning, where a \gls{gan} was trained to generate face features to augment classes with fewer samples~\cite{ding2018one}. \gls{gapf} was proposed to create fair representations of the data in a quantifiable way, allowing for the finding of a de-correlation scheme from the data without access to its statistics~\cite{huang2018generative}. Wang~\etal defined subgroups at a finer-level (\ie Chinese, Japanese, Korean), and determined the familiarity of faces inter-subgroup~\cite{wang2018they}. Genders have also been used to make subgroups (\eg for analysis of gender-based face encodings~\cite{muthukumar2019}). Most recently,~\cite{wang2018racial} proposed adapting domains to bridge the bias gap by knowledge transfer, which was supported by a novel data collection, \gls{rfw}. The release of \gls{rfw} occurred after \gls{bfw} was built - although similar in terms of demographics, \gls{rfw} uses faces from MSCeleb~\cite{guo2016ms} for testing, and assumes CASIA-Face~\cite{yi2014learning} and VGG2~\cite{Cao18} were used to train. In contrast, our \gls{bfw} assumes VGG2 as the test set. 
        % As shown in \cite{wang2018racial}, MSCeleb is highly imbalanced, primarily consisting of images of white individuals. 
        % For this, we expect bias from data, for this is the training set. 
        Furthermore, \gls{bfw} balances gender and race: \gls{rfw} splits subgroups by gender and race, while \gls{bfw} splits by gender, race, or gender and race). 
        % Thus, \gls{rfw} and \gls{bfw} complement one another. % - both add demographic for subjects with faces in renowned, large-scale \gls{fr} datasets.

Most similar to us is~\cite{das2018, demogPairs, lopez2019dataset, srinivas2019face} - each were motivated by insufficient paired data for studying bias in \gls{fr}; then, problems were addressed using labeled data from existing image collections. Specifically, Hupont~\etal curated a set of faces based on racial demographics (\ie \gls{a}, \gls{b}, and \gls{w}) called \gls{dp}~\cite{demogPairs}, while~\cite{srinivas2019face} honed in on adults versus children called \gls{itwcc}. Like the proposed \gls{bfw}, both were built by sampling existing databases, but with the addition of tags for the respective subgroups of interest. Besides, the additional data of \gls{bfw} (\ie added an additional subgroup \gls{i}, along with additional subjects with more faces for all subgroups), we also further split subgroups by gender. Furthermore, we focus on the problem of facial verification and the different levels of sensitivity in cosine similarity scores per subgroup.

\subsection{Human bias in \gls{ml}}
Bias is not unique to \gls{ml} - humans are also susceptible to a perceived bias. In fact, biases exist across race, gender, and even age~\cite{10.1007/978-3-030-13469-3_68, bar2006, meissner2001, nicholls2018}. Wang~\etal showed machines surpass human performance in discriminating between Japanese, Chinese, or Korean faces by nearly 150\%~\cite{wang2018they}, as humans just pass random (\ie 38.89\% accuracy).

We expect human bias to skew to their own genders and races. For this, we measure the human perception with face pairs of different subgroups (Section~\ref{subsec:human-assessment}). The results concur with~\cite{wang2018they}, as we also recorded overall averages below random ($<$50\%). %The details on settings and number of submissions per demographics are discussed in Section~\ref{}.

\glsunset{if}\glsunset{im}\glsunset{af}\glsunset{am}\glsunset{bf}\glsunset{bm}\glsunset{wf}\glsunset{wm}
\section{The BFW Benchmark and Dataset}

We now discuss the \gls{bfw} dataset (\ie collection and preparation), along with the evaluation protocols. We conclude this section by reviewing the settings followed to survey bias in humans.

\subsection{The data}
Problems of bias in \gls{fr} motivated us to build \gls{bfw}. The data evenly represent various subgroups partitioned by demographic. Inspired by \gls{dp}~\cite{demogPairs}, the specification of \gls{bfw} follows in suit, but with additional subgroups (\ie \gls{if} and \gls{im}), an increase in the number of subjects per subgroup, and many more pairs (Table~\ref{tab:ethnic-splits}).

\vspace{1mm}
\noindent\textbf{Compiling subject list.} 
Subjects were sampled from VGG2~\cite{Cao18} - unlike others that depend on multiple sources, \gls{bfw} has fewer potential conflicts in train and test overlap with existing models. We used pre-trained ethnicity~\cite{ambekar2009name} and gender~\cite{levi2015age} classifiers to find candidates for the different subgroups.

\vspace{1mm}
\noindent\textbf{Detecting faces.} Faces were detected using MTCNN~\cite{zhang2016joint}.\footnote{\href{https://github.com/polarisZhao/mtcnn-pytorch}{https://github.com/polarisZhao/mtcnn-pytorch}} Then, assigned into one of two sets. Faces within detected bounding box (BB) regions extended out 130\% in each direction, with zero-padding as the boundary condition made-up one set. The second set were faces aligned and cropped for Sphereface~\cite{liu2017sphereface} (see the next step). Also, coordinates of the BB and the five landmarks from \gls{mtcnn} were stored as part of the static, raw data. For samples with multiple face detections, we used the BB area times the confidence score of the \gls{mtcnn} to determine the face most likely to be the subject of interest, with the others set aside and labeled \textit{miss-detection}.

\begin{figure}[!t] 
	\centering    
 \glsunset{fpr}
  \glsunset{fnr}
	\includegraphics[width=\linewidth]{figures/detcurve-improved.pdf}
		\caption{\textbf{\gls{det} curves.} \emph{Top-left}: per gender. \emph{Top-right}: per ethnicity. \emph{Bottom}: per subgroup (\ie combined). A dashed line shows a difference by about 2$\times$ in \gls{fpr} for the same threshold $\theta_{const}$. \gls{fnr} is the match error count (closer to the bottom is better).}
		\glsreset{det}\glsreset{fpr}
\label{fig:detcurves} 
\end{figure}

\vspace{1mm}
\noindent\textbf{Validating labels.} 
Faces of \gls{bfw} were encoded using the original implementation of the \gls{soa} Sphereface~\cite{liu2017sphereface}: applying an affine transformation to align faces according to predefined eye locations, each face fed through twice (\ie the original and horizontally flipped), with the final encoding obtained by fusing the two features by average pooling (\ie 512 D). A matrix of cosine similarity scores was then generated for each subject and removed samples (\ie rows) with median scores below threshold $\theta=0.2$ (set manually). Mathematically, the $n^{th}$ sample for the $j^{th}$ subject with $N_j$ faces was removed if the ordinal rank of its score $n = \frac{P\times N}{100}\geq\theta$, where $P=50$. In other words, the median (\ie 50 percentile) of all scores for a faces with respect to all of faces for the respective subject must pass a threshold of $\theta=0.2$; else, the face is dropped. This allowed us to quickly prune \gls{fp} face detections. Following~\cite{robinson2016families, robinson2018visual}, we built a JAVA tool to visually validate the remaining faces. For this, the faces were ordered by decreasing confidence, with confidence set as the average score, and then displayed as image icons on top toggling buttons arranged as a grid in a sliding pane window. Labeling then consisted of going subject-by-subject and flagging faces of \emph{imposters}.

\vspace{1mm}
\noindent\textbf{Sampling faces and creating folds.} We created lists of pairs in five-folds with subjects split evenly per subject and without overlap across folds. Furthermore, a balance in the number of faces per was obtained by sampling twenty-five faces at random from each. Next, we generated a list of all the face pairs per subject, resulting in $\sum_{l=1}^{L}\sum_{k=1}^{K_d} {N_k \choose 2}$ positive pairs, where the number of faces of all $K_l$ subjects $N_k=25$  for each of the $L$ subgroups (Table~\ref{tab:ethnic-splits}). Next, we assigned subjects to a fold. To preserve balance across folds, we sorted subjects by the number of pairs and then started assigning to alternating folds from the one with the most samples. Note, this left no overlap in identity between folds. Later, a negative set from samples within the same subgroup randomly matched until the count met that of the positive. Finally, we doubled the total count with negative pairs from across subgroups but in the same fold.
\begin{figure}[t!] 
	\glsunset{fpr}
	\glsunset{sdm}
	\centering
	\centering
	\includegraphics[width=1\linewidth]{figures/violinplots.pdf}
		\caption{\textbf{\Gls{sdm} across subgroups.} Scores of \emph{imposters} have medians around 0.3 but with variations in upper percentiles; \emph{genuine} pairs vary in mean and spread (\eg \gls{af} has more area of overlap). A threshold varying across different subgroups yields a constant \gls{fpr}.} \label{fig:detection-model} 
		\glsreset{sdm}
\end{figure}

% \vspace{-5pt}
\subsection{Problem formulation}\label{subsec:pf} 
% \gls{lfw}~\cite{LFWTech}, a commonly used benchmark for ~\gls{fr}, reserves specific train-test face-pair lists. 
\Gls{fv} is the special case of the two-class (\ie boolean) classification. Hence, pairs are labeled as the ``same'' or ``different'' \textit{genuine} pairs (\ie \textit{match}) or \textit{imposter} (\ie \textit{mismatch}), respectively. This formulation (\ie \gls{fv}) is highly practical for applications like access control, re-identification, and surveillance. Typically, training a separate model for each unique subject is infeasible. Firstly, the computational costs compound as the number of subjects increase.  Secondly, such a scheme would require model retraining each time a new person is added. Instead, we train models to encode facial images in a feature space that captures the uniqueness of a face, to then determine the outcome based on the output of a scoring (or distance) function. Formally put:
\begin{equation}\label{eg:matcher}
    f_{boolean}(\vec{x}_i, \vec{x}_j) = d(\vec{x}_i, \vec{x}_j) \leq \theta
\end{equation}

where $f_{boolean}$ is the \textit{matcher} of the feature vector $\vec{x}$ for the $i^{th}$ and $j^{th}$ sample~\cite{LFWTech}.

Cosine similarity is used as the \emph{matcher} in Eq~\ref{eg:matcher} the closeness of $i^{th}$ and $j^{th}$ features, \ie
$
s_l= \frac{f_i\cdot f_j}{||f_i||_2||f_j||_2}
$ is the closeness of the $l^{th}$ pair.

\subsection{Human assessment}\label{subsec:human-assessment}
We evaluated human on face pairs focusing on two racial groups: Chinese and Caucasians. To focus on the experiment, we honed-in on two groups, white Americans (W) and Chinese from China (C). The purpose was to minimize variability by only analyzing the subsets of the broader groups of whites and Asians.

Samples were collected by recruiting subjects from multiple sources (\eg social media, email lists, and family/friends)-- a total of 120 participants were sampled at random from all the submissions that were (1) complete and (2) from a W or C participant. Specifically, there were 60 W and 60 C, both with \gls{m} and \gls{f} split evenly. A total of 50 face pairs of non-famous ``look-alikes'' were collected from the internet, with 20 ({\emph WA}) and 20 ({\emph C}) pairs (male and female split evenly). The other 10 pairs are of others (\eg Hispanic/ Latino, Japanese, African). The survey was created, distributed, and recorded via \href{https://paperform.co}{PaperForm}.

\begin{figure}[t!]
	\centering    
	\includegraphics[width=.85\linewidth]{figures/confusion.pdf}
		\caption{\small{\textbf{Confusion matrix.} Percent error (Rank 1) for all faces of \gls{bfw} queried against all others. Notice errors concentrate within a subgroup, consistent with the \gls{sdm} in Fig.~\ref{fig:detection-model} (\ie \gls{af} show worst performance, and is mostly confused with faces of the same demographic). This plot is evidence that while race/ethnicity may be challenging to define, the subgroups are meaningful.}}
		\label{fig:confusion} 
\end{figure}

\section{Results and Analysis}
A single \gls{cnn} was used as a means to control the experiments. For this, Sphereface~\cite{liu2017sphereface} trained on CASIA-Web~\cite{yi2014learning}, and evaluated on \gls{lfw}~\cite{LFWTech} (\%99.22 accuracy), encoded all of the faces.\footnote{\href{$https://github.com/clcarwin/sphereface\_pytorch$}{https://github.com/clcarwin/sphereface\_pytorch}} As reported in~\cite{wang2019racial}, \gls{lfw} has about a 13\%, 14\%, 3\%, and 70\% ratio in Asians, Blacks, Indians, and Whites, respectively; furthermore, CASIA-Web is even more unbalanced (again, as reported in~\cite{wang2019racial}), with about  3\%, 11\%, 2\%, and 85\% for the same subgroups.

% Specifically, we analyze three \glspl{cnn} variants: we encode faces using VGG-16~\cite{simonyan2014very},  50-layer residual network (\ie ResNet-50)~\cite{he2016deep}, and \gls{senet} (\ie SENet-50)~\cite{hu2018squeeze}, with results of only the latter (\ie the best performing \gls{senet}) model used throughout the main paper. Results for the other two, alongside \gls{senet} for comparison, are in the Supplemental Material.

\glsunset{am}\glsunset{af}\glsunset{bm}\glsunset{bf}\glsunset{im}\glsunset{if}\glsunset{wm}\glsunset{wf}

\subsection{Score analysis}
Fig.~\ref{fig:detection-model} shows score distributions for faces of the same (\ie \emph{Genuine}) and different (\ie \emph{Imposter}) identity, with a subgroup per \gls{sdm} plot. Notice that score distributions for imposters tend to peak about zero for all subgroups, and with minimal deviation comparing modes of the different plots. On the other hand, the score distribution of the \emph{genuine} pairs varies across subgroups in location (\ie score value) and spread (\ie overall shape). Asian Fem
Fig.~\ref{fig:confusion} shows the confusion matrix of the subgroups. A vast majority of errors occur in the intra-subgroup. It is interesting to note that while the definition of  each group  based on ethnicity and race may not be crisply defined, the confusion matrix indicates that in practice the \gls{cnn} finds that the groups are effectively separate. The categories are, therefore, meaningful in the context of \gls{fr}.

\subsection{\gls{det} analysis}
\glsunset{m}
\glsunset{f}

\begin{figure*}[h!]
    \centering
    \begin{subfigure}[t]{.3\linewidth}
    \includegraphics[width=.75\linewidth]{figures/curve_vgg16_subgroups-crop.pdf}
    \caption{VGG16~\cite{simonyan2014very}}
 \end{subfigure}
    \begin{subfigure}[t]{.27\linewidth}
    \includegraphics[width=.8\linewidth]{figures/curve_resnet50_subgroups-crop.pdf}
    \caption{ResNet50~\cite{he2016deep}}
   \end{subfigure}
    \begin{subfigure}[t]{.27\linewidth}
    \includegraphics[width=.8\linewidth]{figures/curve_senet50_subgroups-crop.pdf}
    \caption{SENet~\cite{hu2018squeeze}}
    \end{subfigure}
    \caption{\textbf{\gls{det} curves for different CNNs}. \gls{fnr} (\%) (vertical) vs \gls{fpr}  (horizontal, log-scale) for VGG2~\cite{Cao18} models with different backbones (VGG16, Resnet50, SEnet50). Lower is better. For each plot, \gls{wm} is the top-performer, while \gls{af} is the worst. The the ordering of the curves is roughly the same for each backbone.}\label{fig:sdm-appendix-a}
\end{figure*}

\gls{det} curves averaged across 5-folds show per-subgroup trade-offs (Fig.~\ref{fig:detcurves}). Note that \gls{m} performs better than \gls{f}, precisely as one would expect from the tails of score-distributions for \emph{genuine} pairs (Fig.~\ref{fig:detection-model}). \Gls{af} and \gls{if} perform the worst.

\begin{table}[b!]
\glsunset{tar}
\glsunset{far}
\caption{\small{\textbf{\gls{tar} at intervals of \gls{far}}. For each subgroup, and overall average \gls{far}, listed are the \gls{tar} scores for a global threshold (top) and the proposed category-based threshold (bottom). Higher is better. The proposed shows improvement in all cases.}}\label{tab:ethnicy-far} 
%  \vspace{-2mm}
\begin{center}
\scriptsize
\begin{tabular}{l c c c c c}
     \gls{far} & 0.3 & 0.1 & 0.01 & 0.001 & 0.0001\\\midrule
    \multirow{2}{.1mm}{\textbf{\gls{af}}} &0.990 & 0.867 & 0.516 & 0.470 & 0.465\\[-4pt]
        &1.000 & 0.882 & 0.524 & 0.478 & 0.474\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{am}}} &0.994 & 0.883 & 0.529 & 0.482 & 0.477\\[-4pt]
        &1.000 & 0.890 & 0.533 & 0.486 & 0.482\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{bf}}} &0.991 & 0.870 & 0.524 & 0.479 & 0.473\\[-4pt]
        &1.000 & 0.879 & 0.530 & 0.484 & 0.480\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{bm}}} &0.992 & 0.881 & 0.526 & 0.480 & 0.474\\[-4pt]
        &1.000 & 0.891 & 0.532 & 0.485 & 0.480\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{if}}} &0.996 & 0.881 & 0.532 & 0.486 & 0.481\\[-4pt]
        &1.000 & 0.884 & 0.534 & 0.488 & 0.484\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{im}}} &0.997 & 0.895 & 0.533 & 0.485 & 0.479\\[-4pt]
        &1.000 & 0.898 & 0.535 & 0.486 & 0.481\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{wf}}} &0.988 & 0.878 & 0.517 & 0.469 & 0.464\\[-4pt]
        &1.000 & 0.894 & 0.526 & 0.478 & 0.474\\[-1pt]
    \multirow{2}{3mm}{\textbf{\gls{wm}}} &0.989 & 0.896 & 0.527 & 0.476 & 0.470\\[-4pt]
        &1.000 & 0.910 & 0.535 & 0.483 & 0.478\\[-1pt]
    \midrule
    \multirow{2}{3mm}{\textbf{Avg.}} &0.992 & 0.881 & 0.526 & 0.478 & 0.474\\[-4pt]
        &1.000 & 0.891 & 0.531 & 0.483 & 0.479\\[-10pt]
\end{tabular}
\end{center}
\glsreset{tar}
\glsreset{far}
%  \vspace{-3mm}

\end{table}

>>>>

% \usepackage{FG2019}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage[acronym]{glossaries}
\usepackage{acronym}
\usepackage{collcell}
\newcommand{\minitab}[2][c]{\begin{tabular}{#1}#2\end{tabular}}

@@

% \let\vec\mathbf
% \usepackage[T1]{fontenc}
% \usepackage[latin9]{inputenc}

\def\colorModel{hsb} %You can use rgb or hsb
\usepackage{babel}
% \usepackage[table]{xcolor}
% 
% \def\colorModel{hsb} %You can use rgb or hsb

\usepackage{hhline}
\newcommand\ColCell[1]{
  \pgfmathparse{#1<50?1:0}  %Threshold for changing the font color into the cells
    \ifnum\pgfmathresult=0\relax\color{white}\fi
  \pgfmathsetmacro\compA{0}      %Component R or H
  \pgfmathsetmacro\compB{#1/100} %Component G or S
  \pgfmathsetmacro\compC{1}      %Component B or B
  \edef\x{\noexpand\centering\noexpand\cellcolor[\colorModel]{\compA,\compB,\compC}}\x #1
  } 
\newcolumntype{E}{>{\collectcell\ColCell}m{0.4cm}<{\endcollectcell}}  %Cell width
\newcommand*\rot{\rotatebox{90}}

\cvprfinalcopy % *** Uncomment this line for the final submission

% \def\cvprPaperID{0001} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

@@

%use this in case of several affiliations
\author{\parbox{16cm}{\centering
    {\large Joseph P Robinson$^1$, Gennady Livitz$^2$, Yann Henon$^2$, Can Qin$^1$,\\ Yun Fu$^1$, and Samson Timoner$^2$}\\
    {\normalsize
    \hspace{-.4in}$^{1}$Northeastern University\hspace{.7in} $^{2}$ISM Connect}}
    % \thanks{}% <-this % stops a space
}

@@

%%%%%%%%% ABSTRACT
\begin{abstract}
We reveal critical insights into problems of bias in state-of-the-art \gls{fr} systems using a novel \gls{bfw} dataset: data balanced for gender and ethnic groups. We show variations in the optimal scoring threshold for face-pairs across different subgroups. Thus, the conventional approach of learning a  global threshold for all pairs resulting in performance gaps among subgroups. By learning subgroup-specific thresholds, we not only mitigate problems in performance gaps but also show a notable boost in the overall performance. Furthermore, we do a human evaluation to measure the bias in humans, which supports the hypothesis that such a bias exists in human perception. For the \gls{bfw} database, source code, and more, visit \href{https://github.com/visionjo/facerec-bias-bfw}{https://github.com/visionjo/facerec-bias-bfw}.
% \href{https://github.com/visionjo/facerec-bias-bfw}{github.com/visionjo/facerec-bias-bfw}
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \begin{bibunit}
\input{sections/introduction}
\input{sections/relatedworks}

\begin{figure}[t!]
    \centering
    \includegraphics[width=.8\linewidth]{figures/montage.pdf}
    \caption{\small{\textbf{\gls{bfw}.} Average face of the different subsets: \emph{top-left}: the entire \gls{bfw}; \emph{top-row} per race;  \emph{left-column}: per gender. The others represent the ethnicity and gender of the race and gender, respectively. Table~\ref{tab:ethnic-splits} defines the acronyms of subgroups.}}
    \label{fig:avg-faces}
\end{figure}

\glsunset{if}\glsunset{im}\glsunset{af}\glsunset{am}\glsunset{bf}\glsunset{bm}\glsunset{wf}\glsunset{wm}
\input{sections/section3}

\begin{figure}[t!]
       \centering
    \includegraphics[width=.95\linewidth]{figures/adaptive_fpr_percent_diff_complete.pdf}
    \caption{\small{\textbf{Percent difference from intended \gls{fpr}.} \emph{Top:} $t_g$ yields \gls{fpr} the span as large as 2$\times$ (\ie 200\%) that intended (\ie \gls{wm} for 1e-4). Furthermore, \gls{f} subgroups tend to perform worse than intended for all cases (while \gls{m} tend to overshoot intended performance, with exception of \gls{im} in for \gls{fpr}=1e-4). \emph{Bottom:} Subgroup-specific thresholds reduces this difference to near zero, where there are small differences, the percent difference across different subgroups is fair (\ie FPR=1e-4).}}\label{fig:percent:difference}
\end{figure}

\input{sections/analysis}
